%%%% necessary customization for the appendix and headers
\appendix 
\makeatletter
\addtocontents{toc}{\protect\renewcommand\protect\cftchappresnum{\@chapapp\ }}
\makeatother
\renewcommand{\thechapter}{\Alph{chapter}}

\chapter{Additional Details} \label{appendix}

%% add your chapter text here
\section{Annotation Instructions}
\label{sec:annotation-instructions}
\subsection{Questions}
\textbf{Question 1 – Which document do you prefer?}
In this question, you are asked to choose which document version you prefer. Some examples of qualities you may use to decide your preference include: 
\begin{itemize}
    \item The quality of the content – The text is grammatical and understandable. E.g. Document A contains major grammatical errors while Document B only contains minor errors.
    \item The formatting – The formatting is reasonable and matches the formatting of the intended document type. E.g. A poster contains panels and each panel contains a header and body text.
    \item The style – The document matches the style of the intended document type. E.g. Shorter, bulleted sentences in a slide deck.
    \item Information represented in the document – The document contains sufficient information to represent the input document. E.g. A blog post represents the most important sections from the input document.
\end{itemize}


The above criteria are non-exhaustive. Not all criteria must be met, and you may use other relevant criteria to make your decision. 
You are not rating the document for factual correctness,\footnote{The annotators are non-experts and do not have the background to determine factual correctness of scientific information. Instead, they are encouraged to use the original paper to understand if the information presented in the documents represent the information in the paper, to the best of their understanding.} and only need to refer to the corresponding scientific article if it will aid in making your preference. 
You can answer this question with either Document A or Document B.


\textbf{Question 2 – On a scale of 1-3, to what degree do you prefer your selection?
}

\noindent In this question you will rate the degree to which you prefer your selection, on the following scale:
\begin{enumerate}
    \item Small preference – The documents are similar in quality and only contain minor differences that affect my preference.
    \item Moderate preference – I clearly prefer one document but the differences are not major.
    \item Strong preference – I have a strong preference for one document and the differences between the documents are major.
\end{enumerate}

\textbf{Question 3 – Why do you prefer your selection? (You may select more than one property)}
\begin{itemize}
    \item[$\square$] Formatting
    \item[$\square$] Information
    \item[$\square$] Quality
    \item[$\square$] Style
    \item[$\square$] Other (free text)
\end{itemize}


\subsection{Edge cases}
For most edge cases, it is up to your discretion on how to best handle the case. However, below are a few examples of how you could consider certain edge cases:

\textbf{Example 1: Slides 1-5 of Document A are higher quality but slides 6-10 of Document B are higher quality.
}
You could reason that the first slides represent the most important information, and choose Document A. However, since Document B contained higher quality slides for another portion of the document, you could rate your degree of preference as “Small preference.”

\textbf{Example 2: Document A more closely matches the style of the intended document type, but Document B contains more relevant information to the source document.
}
You could consider if Document A contains sufficient information to represent the input document, such as representing the most important sections. If yes, then you could prefer Document A. If not, then you could reason that information content is more important than style, and prefer Document B.

\textbf{Example 3: Document A contains more relevant information than Document B, but also contains major formatting errors, such as text being cut off from the document.}

You could reason that although Document A contains more relevant information, the major formatting errors are significant enough to prefer Document B.

\textbf{Example 4: Neither document matches the style or formatting of the intended document type. 
}
Since neither document matches the style or formatting of the intended document type, you could consider other criteria, such as quality of content or the information represented. 


